\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-03-21 paper/main.tex
.

The prediction $P_k$ is not dependent on the drug $d_i$ and will
predict the same ranked list of drugs for all proteins; consequently,
this na\"ive predictor will not predict any useful information about
interactions between drugs and proteins, but allows us to estimate the
effect of potential biases in our evaluation dataset.
% Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.

Our na\"ive predictor is not applicable for a protein split
cross-validation as the number of interactions for each protein in the
validation set is unknown.  We apply this naive predictor on both the
STITCH and Yamanishi datasets, using the full dataset as well as
5-fold cross-validation over drugs as well as over drug--protein
pairs.  For each fold, we gradually increase $k$ to determine the best
performance for each fold. Using the full dataset, drug--target split,
and drug split, we obtain the following MacroAUC results: for the
STITCH database, we obtain a performance of $0.76$ on the whole
dataset, $0.70$ for the drug--target pairs and $0.73$ in case of the
drug splitting scheme; on the Yamanishi dataset, we obtain MacroAUC
scores of $0.88$, $0.84$ and $0.85$, for the total dataset,
drug--target pair and drug split, respectively.  The na\"ive predictor
shows higher performance on the Yamanishi dataset than on STITCH, and
a substantial gain in comparison to an expected random predictor on
both datasets. In the following, we utilize this nai\"ve predictor as
baseline to compare its performance to state of the art models and
\name.

@@

We evaluate all models on their recommended splitting scheme choice,
hyperparameters and folds in cross-validation, measuring their
respective AUROC. We further evaluate each model by performing a
protein-wise cross-validation determining the MacroAUC and
$MicroAUC_p$. For this evaluation, we allow stratification for the
training process but not for the validation and testing phase as real
world applications of these models would have to deal with possibly
imbalanced data.

The results of our experiments are summarized in part (c) of Table
\ref{tab:Results}; we calculated the performance of all compared
methods over their original splitting scheme and over a protein
split. First, we find that there is a large difference in performance
when evaluating over a drug--target pair split compared to a
protein split, with generally higher performance achieved when using
the drug--target pair split.
Second, when evaluating the same methods over a protein split, we find
a substantial performance difference in comparison to the splitting
scheme used in the original evaluation of each method. DTI-CDF was
originally evaluated on all three splitting schemes underlining this
point \cite{}. While \name provides comparable performance to DTI-CDF,
it yields considerably better results with respect to $MicroAUC_p$.

As the difference in performance with different splitting schemes is
quite striking, we evaluated additional drug--target interaction and
drug--target affinity prediction methods that were trained and
evaluation on other datasets.  Following the results of MolTrans
(\textbf{CITATION}), we reevaluated DeepDTI (\textbf{CITATION}),
DeepDTA (\textbf{CITATION}), DeepConv-DTI (\textbf{CITATION}) and
MolTrans itself on BioSnap\todo{What is that? Is it in methods?}
(\textbf{CITATION}) over a protein split, as shown in part (d) of
Table \ref{tab:Results}. The authors of MolTrans evaluated all
considered methods over the drug--target pair and the protein split;
we were able to reproduce the results (Table \ref{tab:Results} (d)),
showing a substantial difference based on the splitting scheme. We
additionally computed the $MicroAUC_p$ score for all considered
methods, leading to comparable\todo{are the datasets comparable?}
scores as for the results obtained on the Yamanishi dataset.

>>>>

The prediction $P_k$ is not dependent on the drug $d_i$ and will
predict the same ranked list of drugs for all proteins; consequently,
this na\"ive predictor does not rely on any biological features and
will not predict any novel information about interactions between
drugs and proteins; the na\"ive predictor only exploits imbalances in
the evaluation set to make predictions that may perform
well. % but allows us to estimate the effect of potential
% biases in our evaluation dataset.
% Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.

The way in which we formulated the na\"ive predictor, it is not
applicable for a protein split cross-validation as the number of
interactions for each protein in the validation set is unknown.  We
apply this naive predictor on both the STITCH and Yamanishi datasets,
using the full datasets as well as a 5-fold cross-validation over
drugs and over drug--protein pairs to compare the prediction results
directly to \name.  For each fold, we gradually increase $k$ to
determine the best performance for each fold. Using the full dataset,
drug--target split, and drug split, we obtain the following MacroAUC
results\todo[inline]{Put in Results table, then remove from here}: for
the STITCH database, we obtain a performance of $0.76$ on the whole
dataset, $0.70$ for the drug--target pairs and $0.73$ in case of the
drug splitting scheme; on the Yamanishi dataset, we obtain MacroAUC
scores of $0.88$, $0.84$ and $0.85$, for the total dataset,
drug--target pair and drug split, respectively.  The na\"ive predictor
shows higher performance on the Yamanishi dataset than on STITCH, and
a substantial gain in comparison to an expected random predictor on
both datasets. In the following, we utilize this nai\"ve predictor as
baseline to compare its performance to state of the art models and
\name.

@@

We evaluate all models on their recommended splitting scheme choice,
hyperparameters and folds in cross-validation, measuring their
respective AUROC. We further evaluate each model by performing a
protein-wise cross-validation determining the MacroAUC and
$MicroAUC_p$. For this evaluation, we allow
stratification\todo{Stratification may not be the right word here; do
  you mean ``sub-sampling of negatives''?} for the training process
but not for the validation and testing phase as real world
applications of these models would have to deal with possibly
imbalanced data.

The results of our experiments are summarized in part (c) of Table
\ref{tab:Results}\todo{Check after restructure}; we calculated the
performance of all compared methods over their original splitting
scheme and over a protein split. We find that there is a large
difference in performance when evaluating over a drug--target pair
split compared to a protein split, with generally higher performance
achieved when using the drug--target pair split.  Second, when
evaluating the same methods over a protein split, we find a
substantial performance difference in comparison to the splitting
scheme used in the original evaluation of each method. DTI-CDF was
originally evaluated on all three splitting schemes underlining this
point \cite{}. While \name provides comparable performance to the
na\"ive predictor and DTI-CDF in terms of MacroAUC, it yields
considerably better results with respect to $\textrm{MicroAUC}_p$. We
also find that methods that are trained using a protein split
generally result in higher $\textrm{MicroAUC}_p$ than methods trained
using a drug--target pair split.

As the difference in performance with different splitting schemes is
quite large, we further evaluated additional drug--target interaction
and drug--target affinity prediction methods that were trained and
evaluation on other datasets.  Following the results of MolTrans
\cite{}\todo{add}, we reevaluated DeepDTI \cite{}\todo{add}, DeepDTA
\cite{}, DeepConv-DTI \cite{} and MolTrans itself on BioSnap\todo{What
  is that? Is it described in methods?}  \cite{} over a protein split,
as shown in part (d) of Table \ref{tab:Results}\todo{check table}. The
authors of MolTrans evaluated all considered methods over the
drug--target pair and the protein split; we were able to reproduce the
results (Table \ref{tab:Results} (d))\todo{check table}, showing a
substantial difference based on the splitting scheme. We additionally
computed the $MicroAUC_p$ score for all considered methods, leading to
comparable\todo{are the datasets comparable?} scores as for the
results obtained on the Yamanishi dataset.
